---
title: "Homework4"
format: html
editor: visual
editor_options: 
  chunk_output_type: inline
---

## Task 1: Conceptual Questions

1.  If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?

> "../MyData.csv"

2.  What are the major benefits of using R projects?

> An R project sets up and uses a directory structure that makes it easy to reference files relatively, as well as generate PDFs and html files, and rendering .qmd files.

3.  What is git and what is github?

> Git is a software package that provides version control for confined repositories. Github is a user-friendly way to access git capabilities, since it provides a nice user interface with which to utilize git version control.

4.  What are the two main differences between a tibble and a data.frame?

> First, a tibble has a prettier printing method, so it is displayed nicely when printed. Second, when a column of a tibble is selected using brackets and indices (e.g. tibble[rows, cols]), the returned object is a tibble rather than a vector.

5.  Rewrite the following nested function call using BaseR’s chaining operator: arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length \< 1.55), Species)

>| iris |>
 |      as_tibble() |>
 |      select(starts_with("Petal"), Species) |>
 |      filter(Petal.Length \< 1.55) |>
 |      arrange(Species)


## Task 2: Reading Delimited Data

### Glass Data

Read data from URL (data is comma delimited):

```{r}
library(readr)

glass_data <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data", 
                       col_names=c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type of Glass"))

glass_data
```

Overwrite "Type_of_glass" column values and filter data:

```{r, warning=FALSE,message=FALSE}
library(dplyr)

glass_data <- glass_data |>
  mutate(`Type of Glass` = case_when(
    `Type of Glass` == 1 ~ 'building_windows_float_processed',
    `Type of Glass` == 2 ~ 'building_windows_non_float_processed',
    `Type of Glass` == 3 ~ 'vehicle_windows_float_processed',
    `Type of Glass` == 4 ~ 'vehicle_windows_non_float_processed',
    `Type of Glass` == 5 ~ 'containers',
    `Type of Glass` == 6 ~ 'tableware',
    `Type of Glass` == 7 ~ 'headlamps'
  )) |>
  filter(Fe < 0.2 & (`Type of Glass` %in% c('tableware', 'headlamps') ))

glass_data
```

### Yeast Data

Read data from URL (data is fixed width):

```{r}
yeast_data <- read_fwf("https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
                       fwf_widths(c(12, 6, 6, 6, 6, 6, 6, 6, 6, 3),
                          c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class")
                       )
)
yeast_data
```

Remove seq_name and nuc columns, add new columns:

```{r}
yeast_data |>
  select(-c(seq_name, nuc)) |>
  group_by(class) |>
  mutate(across(where(is.numeric), list(mean=mean, median=median), .names="{.col}_{.fn}"))
```

## Task 3: Combining Excel and Delimited Data

Import white wine data:

```{r}
library(readxl)
white_wine <- readxl::read_excel("white-wine.xlsx")

white_wine
```

Rename columns:

```{r}
white_wine_ancil <- readxl::read_excel("white-wine.xlsx",
                                       sheet = excel_sheets("white-wine.xlsx")[2])

names <- pull(white_wine_ancil, 1)

names

names(white_wine) <- names

white_wine

```

Add type column:
```{r}
white_wine <- white_wine |>
                mutate(type="white")

white_wine
```

Import red wine data, replace column names, create type column:

```{r}
red_wine <- readr::read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", delim = ";")

red_wine

names(red_wine) <- names

red_wine <- red_wine |>
                mutate(type="red")

red_wine
```

Combine datasets:
```{r}
wine_data <- dplyr::bind_rows(white_wine, red_wine)

wine_data
```

Filter, sort, and change the columns of the data:

```{r}
wine_data |>
  filter(quality>6.5 & alcohol<132) |>
  arrange(desc(quality)) |>
  select(contains("acid"), alcohol, type, quality) |>
  group_by(quality) |>
  mutate(alcohol_mean=mean(alcohol), alcohol_sd=sd(alcohol))
```
